---
title: "Exploratory Data Analysis-Project 2"
author: "WNH"
date: "January 2, 2020"
output: html_document

---
The overall goal of this assignment: to explore the National Emissions Inventory database and see what it say about fine particulate matter pollution in the United states over the 10-year period 1999-2008. You my use any R package you want to support your analysis.

```{r}
library("ggplot2")
library("lattice")
library("dplyr")
```

```{r}
fileurl <- "https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2FNEI_data.zip"
SCC_PM25<-download.file(fileurl, "./exdata_data_NEI_data.zip")
if(!file.exists("exdata_data_NEI_data")) {
  unzip("exdata_data_NEI_data.zip")
}

dir()
```

```{r}
NEI<-readRDS("summarySCC_PM25.rds",refhook=NULL)
str(NEI)
dim(NEI)
```

```{r}
SCC<-readRDS("Source_Classification_Code.rds")
str(SCC)
```
# Question 1: Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? Using the  base plotting system, make a plot showing the total PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.
```{r}

PM25Year<- NEI %>%group_by(year)
totalPM25 <- PM25Year %>% summarise(
  Emissions=sum(Emissions)
) 
totalPM25
str(totalPM25)
barplot <- barplot(height=totalPM25$Emissions/1000,names.arg=totalPM25$year, ylim=c(0,8000),xlab="Year",
        ylab="Total PM2.5 Emissions(kiloton)", main=("Total PM2.5(Kiloton) Emissions Over Years"))

text(x=barplot,y=round(totalPM25$Emissions/1000,2), labels=round(totalPM25$Emissions/1000,2),pos = 3, cex = 1, col = "black")

```
#method2, total pm2.5 emissions over years

```{r}
totalPM25byYear <- NEI %>%
  group_by(year) %>%
  summarise(Emissions=sum(Emissions))
with(totalPM25byYear,barplot(
  height=Emissions/1000,
  names.arg=year,
  ylim=c(0,8000),
  xlab="Year",
  ylab="Total PM2.5 Emissions(kiloton)",
  main="Total PM2.5 Emissions Over Years(1999 - 2008)",
  col = grey.colors(4)
))

text(x=barplot,y=round(totalPM25byYear$Emissions/1000,2), labels=round(totalPM25byYear$Emissions/1000,2),pos = 3, cex = 1, col = "black")

```
# Question2:Have total emissions from PM2.5 decreased in the Baltimore City, Maryland (fips == "24510") from 1999 to 2008? Use the base plotting system to make a plot answering this question.
```{r}
totalPM25byCity <- NEI %>%
  group_by(fips,year) %>%
  filter(fips=="24510") %>%
  summarise(Emissions=sum(Emissions))
with(totalPM25byCity,barplot(
  height=Emissions/1000,
  names.arg=year,
  ylim=(c(0,4)),
  xlab="Year",
  ylab="Total PM2.5 Emissions in Baltimore,MD(kilotons)",
  main="Total PM2.5 Emissions in Baltimore,MD Over Years(1999 - 2008)",
  col = grey.colors(4)
))

text(x=barplot,y=round(totalPM25byCity$Emissions/1000,2), labels=round(totalPM25byCity$Emissions/1000,2),pos = 3, cex = 1, col = "black")
```
# 3.Of the four types of sources indicated by the type (point, nonpoint, onroad, nonroad) variable, which of these four sources have seen decreases in emissions from 1999-2008 for Baltimore City? Which have seen increases in emissions from 1999-2008? Use the ggplot2 plotting system to make a plot answer this question.
```{r}
PM25byBaltimore <- NEI %>% 
  select(year,fips,type,Emissions)%>%
  filter(fips=="24510") %>%
  group_by(year,type)

g<-ggplot(PM25byBaltimore,aes(x=factor(year),y=Emissions,fill=type))
g+geom_bar(stat="identity")+
  labs(title="PM2.5 Emissions in Baltimore 1999-2008 by Source Types")+
  labs(x="Year", y="PM2.5 Emissions in Baltimore")+
  geom_smooth(method="lm")+
  facet_grid(.~type)+
  scale_fill_grey()+
  theme_bw()




```
# 4. Across the United States, how have emissions from coal combustion-related sources changed from 1999-2008?
```{r}
SCC_Coal <- grepl("Fuel Comb.*Coal", SCC$EI.Sector)
SCC_Coal_Source <- SCC[SCC_Coal,]

temp <- NEI[(NEI$SCC %in% SCC_Coal_Source$SCC),]
PM25byCoal <- summarise(group_by(temp, year), Emissions=sum(Emissions))

ggplot(PM25byCoal, aes(x=factor(year), y=Emissions/1000,fill=year, label = round(Emissions/1000,2))) +
  geom_bar(stat="identity") +
  xlab("year") +
  ylab(expression("Total PM2.5 Emissions(kilotons)")) +
  ggtitle("Emissions(Kilotons) from Coal Combustion-related Sources in U.S")+
  geom_label(aes(fill = year),colour = "gray")

```
#5. How have emissions from motor vehicle sources changed from 1999-2008 in Baltimore City?

```{r}
vehicles <- grepl(pattern = "Vehicle",x = SCC$SCC.Level.Three,ignore.case = TRUE)
vehiclesSCC <- SCC[vehicles,]$SCC
vehiclesNEI <- NEI[NEI$SCC %in% vehiclesSCC,]

Baltimorevehiclesnei <- (vehiclesNEI[vehiclesNEI$fips == "24510",])

g <- ggplot(data = Baltimorevehiclesnei, aes(factor(year), Emissions)) 
        
g+ geom_bar(stat = "identity",fill = "grey", width = 0.75) +
        scale_fill_grey()+
        theme_bw() +
        labs(x="Year", y=expression("Total PM2.5Emission (Tons)")) + 
        labs(title=expression("PM2.5 Vehicle Source Emissions, Baltimore City"))


```

#6.Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in Los Angeles County, California (fips=="06037"). Which city has seen greater changes over time in motor vehicle emissions?
```{r}
VehicleSCC <- grepl(pattern="Vehicle",x=SCC$SCC.Level.Three,ignore.case = TRUE)
SCCVehicle <- SCC[VehicleSCC,]$SCC
NEIVehicleSCC <- NEI[NEI$SCC %in% SCCVehicle,]
BaltimorevehiclesNEI <- subset(NEIVehicleSCC,fips=="24510")
BaltimorevehiclesNEI$city <- "Baltimore City"
LosAnglesvehiclesNEI <- subset(NEIVehicleSCC, fips=="06037")
LosAnglesvehiclesNEI$city <- "Los Angeles County"
EmissionMotorVehicle <- rbind(BaltimorevehiclesNEI,LosAnglesvehiclesNEI)

g <- ggplot(data=EmissionMotorVehicle,aes(factor(year), fill= city,Emissions))
g + geom_bar(stat="identity")+
  labs(title="Comparison of PM2.5 Emissions from Motor Vehicle Sources between Baltimore City and Los Angeles County")+
  labs(x="Year", y="PM2.5 Emissions")+
  facet_grid(.~city)+
  scale_fill_grey()
  
```

```{r}

```



```{r}

```

